---
title: "Customer churn analysis"
author: "KSD"
date: "January 10, 2018"
output: 
    # pdf_document: default
    html_document: default
    toc: true
---

```{r}
library(knitr)
opts_chunk$set(tidy = TRUE)
```

##Loading libraries
```{r}
library(ggplot2)
library(dplyr)
library(reshape2)
```
## Loading data
```{r}
d <- read.csv('data.csv',header=T)
d[,1] <- as.character(d[,1])
#d[,2] <- as.numeric(d[,2])
str(d)
```
## Data manipulation
```{r}
t <- c(1:length(d[,2])) # t= month number
d1 <- cbind(d,t)
d1 <- mutate(d1,S=Active,t1=t-1) # t1 = preiod number of each month; S = active customers;
d1 <- mutate(d1, N= -c(NA,diff(d1$S))) # N = churned customers
d2 <- d1 %>% select(t1, S, N)
S0 <-  d2[["S"]][1] #initial cohort size
N1 <- d2[["N"]][-1] # vector of incremental churns
S1 <- d2[["S"]][length(t)] #last element of active customers column
```

# Q1 BG model
## Defining the log likelihood of BG
```{r}
# log of P
log_P <- function(t, a, b) {
  lbeta(a+1, b+t-1) - lbeta(a, b)
}
# log of S
log_S <- function(t, a, b) {
  lbeta(a, b+t) - lbeta(a,b)
}
#log likelihood of Beta Geometric functions 
LL_BG <- function(pars, N, S) {
  a <- exp(pars[1])
  b <- exp(pars[2])
  Tobs <- length(N)
  t <- 1:Tobs
  LL_p <- sum(N * log_P(t,a, b))
  LL_s <- S * log_S(Tobs, a, b)
  LL_all <- LL_p + LL_s
  return(-LL_all) ## scalar
}
```
## Optimizing a and b parameters of the Beta function
```{r}
opt_BG <- optim(c(0,0),fn=LL_BG,
N=N1, S=S1)
a <- exp(opt_BG[["par"]][1])
b <- exp(opt_BG[["par"]][2])
LL <- -opt_BG[["value"]]
#Log likelihood estimates of a and b are:
print(a);print(b)
#Log likelihood value at the estiamted parameters:
print(LL)
```

# Q2
##(a)Probability that this customer will cancel service after only one month
```{r}
exp(log_P(1,a,b)) #t=1. P(T=1|a,b)=(theta*(1-theta)^2)
```

##(b)Probability that this customer will cancel service after 2 months
```{r}
exp(log_P(2,a,b)) #t=2. P(T=2|a,b)=(theta*(1-theta)^2)
```
##(c)Probability that he will renew for March
```{r}
exp(log_S(3,a,b)) #t=3 for March. S(3)=(1-theta)^3
```
##(d)renewal probability for a customer who remained active through August
```{r}
exp(log_S(8,a,b)) #t=8 for August. S(8)=(1-theta)^3
```
##(e)Members active through the end of the year
```{r}
#To forecast customer counts, evaluate the survival function at each value of ???, and multiply by the size of the cohort(ref: Week_1_slides,pg34)
round(S0*exp(log_S(12,a,b)),0) #t=12 for December. S(3)=(1-theta)^3
```

# Q3
##(a)

```{r}
# Making the table shown in Q3(a)
S3 <- matrix(0,nrow=12, ncol=4)
for (i in c(1:4)){
  S3[i,i]=2132
}
```

```{r}
S0 <- 2132 #intial cohort size
t2 <- c(1:12) #Sep:august
S_pred2 <- array() #declaring an empty array to be used below
S4 <- S3

for (i in c(1:4)){
  t1 <- 12-i
  t <- c(1:t1)
  S01 <- S3[i,i] #2132 new customers
  S_pred2 <- round(S01*exp(log_S(t,a,b)),0) #array of predicted remaining customers 
  S4[,i] = c(S3[1:i,i], S_pred2) #all cohorts put into one matrix
}
rownames(S4) <- c(month.name[9:12],month.name[1:8])
colnames(S4) <- month.name[c(9:12)]
total <- rowSums(S4) 
S4 <- cbind(S4, total)
#S4 description: Col1 - The remainder from January cohort; col2 - The predicted customers for Sept...similarly the following columns; last column contains row sums
```

##(b) Retention rates
## Predicted counts from BG models
```{r}
# calcualting the customers active after each interval
S0 <- d[[2]][1] #intial cohort size
t2 <- c(0:8) #interval for the year and for next year until August
S_pred=round(S0*exp(log_S(t2,a,b)),0)
```

```{r}
d4 <- cbind('January'= c(S_pred, rep(NA,dim(S4)[1]-length(S_pred))), S4) #combining the BG coutns from Jan to the other cohorts
reten_rate <- apply(d4, 2, function(x){x/lag(x,1)}) #Retention rate of all cohorts
reten_rate <- data.frame(reten_rate)
reten_rate[reten_rate>1] <- NA # replace retention rate values greater than 1
```

```{r}
#plotting all the retention rate
months <- colnames(reten_rate)
leg <- colnames(reten_rate)
matplot(c(1:12), reten_rate, type = 'l', xlab="Months", ylab="Retention Rate")
```





